% !TEX root = main.tex

% AWR: Guidelines (Unit 7, page 5):
% At most 300 words
% Content:
% Purpose of the study
% Research problems
% Basic design of the study
% Summary of interpretations and conclusions

% AWR: Parts of an abstract (checklist) (Unit 7, page 6):
% Motivation
% Problem statement
% Approach
% Results
% Conclusions

% The state-of-the-art \acrlongpl{atp} for \acrlong{fol}, such as Vampire,
State-of-the-art superposition-based theorem provers for \acrlong{fol} % such as E or Vampire,
% - imho sem patri neurcity clen (the 0 clen v mnoznem cisle)
% - byl specifictejsi z vymezenim katergorie dokazovacu
% - to ``such as'' bych pripadne vynechal
% use \glspl{sot} to constrain the proof search.
rely on \glspl{sot} to constrain the applicability of inference rules 
which in turn shapes the ensuing search space.
The popular Knuth-Bendix simplification ordering is parameterized by 
a \emph{symbol precedence}---a permutation of the predicate and function symbols
% occurring in the input problem.
% maybe:
of the input problem's signature.
The choice of the precedence thus has an indirect yet often substantial impact 
% on the length of the proof search.
on the ammount of work required to successfully complete a proof search.

\todo{MS: idea maybe to work out more in the introduction: 
We could stress how the impact or the choice of a precedence is \emph{indirect}, 
(as it's already obvious from the explanation here)
all the more interesting one can learn from observing just this indirect impact 
which precendences are good and which are bad!}

This paper describes and evaluates a symbol precedence recommender,
a machine learning system which estimates the best possible precedence
based on observations of prover performance on a set of problems and random precedences.
Using the graph convolutional neural network technology % \acrlong{gcn} % how to put neural in the middle?
the system does not presuppose the problems to be related or share a common signature. 
When coupled with the theorem prover Vampire and evaluated on the \acrshort{tptp} problem library
the recommender is found to outperform a state-of-the-art heuristic by more than \SI{2}{\percent}.

% When trained on proof attempts that use random symbol precedences,
% the recommender outperforms the state-of-the-art heuristic by more than \SI{2}{\percent}.
% The recommender is evaluated on \acrlong{fol} problems from the \acrshort{tptp} problem library.
% \todo{FB: Should I mention Vampire here?}
% \todo{FB: Try to mention that the recommender is signature-agnostic.}

\iffalse
This paper describes and evaluates a symbol precedence recommender based on a \acrlong{gcn}.
When trained on proof attempts that use random symbol precedences,
the recommender outperforms the state of the art heuristic by more than \SI{2}{\percent}.
The symbol precedences predicted by the recommender
specify Knuth-Bendix \glspl{sot}.
These orderings in turn guide the proof search in the state-of-the-art \acrlong{atp} Vampire
by constraining the inferences in the superposition calculus.
\todo{MS: uz i ty ``random symbol precedences'', kterymi to vypraveni nahore zacina potrebuji vampira.
Mozna bude lepsi, trochu nudne, zacit tim, odkud se berou a na co se pouzivaji precedence
a teprve pak rict, ze se z nich ucime a pak umime navrhovat nove, lepsi?}
The recommender is evaluated on \acrlong{fol} problems from the \acrshort{tptp} problem library.
\fi

\keywords{saturation-based theorem proving \and
% \gls{sot} \and
simplification ordering \and % isn't it great to be sometimes able to say ``simplification ordering on terms'' and sometimes just "simplification ordering"?
symbol precedence \and
\acrlong{ml} \and
\acrlong{gcn}}

% Motivation and purpose of the study are excluded because the audience at CADE need not be reminded of the motivation for making the provers faster.
